Unifying Lower Bounds on Prediction Dimension of Consistent Convex Surrogates

### 2021-01-27 colt-21/cvx-flats.tex
Pass through the appendix for consistency's sake lol

\section{Conclusions and future work}\label{sec:conclusions}
In this work, we show that indirect property elicitation can be a powerful necessary condition for the existence of a consistent surrogate loss (Theorem~\ref{thm:consistent-implies-indir-elic}).
Furthermore, we introduce a new lower bound (Corollaries~\ref{cor:Pcodim-flat-single-val-prop} and~\ref{cor:Pcodim-flat-elic-relint-prop}) on the dimension of a consistent convex loss that is generally applicable and extends previous results from both the discrete and continuous estimation settings (Corollaries~\ref{cor:fsd-bound}, \ref{cor:variance}, \ref{cor:spectral-risks}).

@@

\newpage
\appendix
\section{A general notion of calibration}\label{app:calibration}
For general settings, we introduce a notion of calibration that is a special case of calibration as introduced by~\cite[Definition 2.7]{steinwart2007compare} and~\citet[Chapter 3]{steinwart2008support}. 
\jessiet{I think we need to be careful how we present this; I'm not sure it's a new definition, though we seem to prove new results about it.}
We will show that in discrete prediction settings, it is equivalent to the more commonplace definition given in Definition~\ref{def:calibrated-finite}.
Therefore, we use this more general definition of calibration when proving statements about the relationship between consistency, calibration, and indirect elicitation.

\begin{definition}[Calibrated]\label{def:calibrated-general}
	A loss $L:\reals^d \times \Y \to \reals$ is \emph{calibrated} with respect to a loss $\ell : \R \times \Y \to \reals$ eliciting the property $\gamma$ \jessie{$\gamma$ isn't used here...?} if there is a link $\psi : \reals^d \to \R$ such that, for all distributions $p \in \P$, there exists a function $\zeta : \reals_+ \to \reals_+$ with $\zeta$ continuous at $0^+$ and $\zeta(0) = 0$ such that for all $u \in \reals^d$, we have
	\begin{equation}\label{eq:calibrated-general}
	\ell( \psi(u); p) - \risk{\ell}(p)  \leq \zeta \left(  \exploss{L}{u}{p} - \risk{L}(p) \right)~.~
	\end{equation}
\end{definition}

@@

\begin{proposition}
	When $\R$ and $\Y$ are finite, a continuous loss and link $(L, \psi)$ are calibrated with respect to a target loss $\ell$ via Definition~\ref{def:calibrated-general} if and only if they are calibrated via Definition~\ref{def:calibrated-finite}.
  %\btw{Okay if $\Gamma$ is empty}
\end{proposition}
\begin{proof}
$\implies$
	We prove the contrapositive; if $(L, \psi)$ is not calibrated with respect to $\ell$ by Definition~\ref{def:calibrated-finite}, then it is not calibrated via Definition~\ref{def:calibrated-general} either.
	If $(L, \psi)$ are not calibrated with respect to $\ell$ by Definition~\ref{def:calibrated-finite}, then there is a $p \in \P$ so that $\inf_{u : \psi(u) \not \in \gamma(p)} \exploss{L}{u}{p} = \inf_u \exploss{L}{u}{p}$.
	Thus there is a sequence $\{u_m\}$ so that $\lim_{m \to \infty} \psi(u_m) \not \in \gamma(p)$ and $\exploss{L}{u_m}{p} \to \risk{L}(p)$.  
	Now we have $R_L(u_m; p) \to 0$ but $R_\ell(\psi(u_m); p) \not \to 0$, so by Lemma~\ref{lem:calib-converging-regrets}, we contradict calibration by Definition~\ref{def:calibrated-general}.
%	
% Commented out 05.19.2020 for easier proof if Lemma 5 is true.
%	We prove the contrapositive; if $(L, \psi)$ is not calibrated with respect to $\ell$ by Definition~\ref{def:calibrated-finite}, then it is not calibrated via Definition~\ref{def:calibrated-general} either.
%	
%	Suppose there was a distribution $p \in \simplex$ so that $\inf_{u : \psi(u) \not \in \gamma(p)} L(u;p) = \inf_{u} L(u;p)$.
%	There must then be a sequence $\{u_m\} \to u$ so that $\lim_{m \to \infty} \psi(u_m) \not \in \gamma(p)$ and $L(u_m; p) \to \risk{L}(p)$.
%	
%	This consequently implies $R_L(u_m;p) \to 0$ as $L(u_m; p) \to \risk{L}(p)$, but as $\ell(\psi(u_m); p) \not \to \risk{\ell}(p)$ (if it did converge, then we would have $\psi(u_m) \to r \in \gamma(p)$), so $R_\ell(\psi(u_m); p) \not \to 0$.
%	Thus, by Lemma~\ref{lem:calib-converging-regrets}, we have no calibration via Definition~\ref{def:calibrated-general}.

$\impliedby$
Suppose there was a function $\zeta$ satisfying the bound in Equation~\eqref{eq:calibrated-general} for a fixed distribution $p \in \P$.
Observe the bound in Equation~\eqref{eq:calibration} can be written as $R_L(u,p) > 0$ for all $p \in \simplex$ and $u$ such that $\psi(u)$ is bounded away from $\gamma(p)$. \jessiet{details correct??}

By Equation~\eqref{eq:calibrated-general}, for any sequence $\{u_m\}$ so that $\psi(u_m) \not \to \gamma(p)$, we have must have $\zeta(R_\ell(\psi(u_m), p)) \not \to 0$ as we would otherwise contradict the bound in Equation~\eqref{eq:calibrated-general} since $R_\ell(\psi(u), p) \not \to 0$. 
Therefore $R_L(u_m, p) \not \to 0$; thus, the strict inequality holds.
\end{proof}

@@

Lemma~\ref{lem:calib-converging-regrets} now gives a condition equivalent to calibration without requiring one to already have a function $\zeta$ in mind.
\begin{lemma}\label{lem:calib-converging-regrets}
	A continuous surrogate and link $(L,\psi)$ are calibrated (via definition~\ref{def:calibrated-general}) with respect to $\ell$ if and only if, for all $p \in \P$ and sequences $\{u_m\}$ so that $R_L(u_m; p) \to 0$, we have $R_\ell(\psi(u_m); p) \to 0$.
	\jessie{$(A \wedge C) \iff D$}
\end{lemma}
\begin{proof}
\jessie{$(A \wedge C) \implies D$}
	$\implies$ Take a sequence $\{u_m\}$ so that $R_L(u_m;p) \to 0$.
	Since $\zeta(0) = 0$ and $\zeta$ is continuous at $0$, we have $\zeta(R_L(u_m;p)) \to 0$.
	As the bound from Equation~\eqref{eq:calibrated-general} is satisfied for all $u \in \reals^d$ by assumption, we observe
	\begin{align*}
	\forall m, \; &0 \leq R_\ell(\psi(u_m); p) \leq \zeta(R_L(u_m;p))\\
	\implies &0 \leq \lim_{m \to \infty} R_\ell(\psi(u_m); p) \leq \lim_{m \to \infty} \zeta(R_L(u_m;p)) = 0\\
	\implies &0 = \lim_{m\to\infty} R_\ell(\psi(u_m); p) ~.~
	\end{align*}

@@

  (ii)
  We first show $\spn(F_p) = \Scr_F(p)$.
  First, take $v \in \Scr_F(p)$, and take $\epsilon_0$ as in the definition.
  For $\epsilon = \epsilon_0 / 2$, we then have $p + \epsilon v \in F \implies \epsilon v \in F_p$, and therefore, $v \in \spn(F_p)$.  
  Now take $v \in \spn(F_p)$.
  Since $p \in \relint{F}$ (i), we have $\vec 0 \in \relint{F_p}$.
  Therefore there is an $\epsilon_0 > 0$ so that $\epsilon v \in F_p$ for all $\epsilon \in (-\epsilon_0, \epsilon_0)$ by convexity of $F$.
  Therefore, $v \in \Scr_F(p)$, and the equality follows.

@@

	Taking $\P = \simplexp$, we know $p' \in \relint{\simplexp}$ by construction, so we can apply Lemma~\ref{lem:finite-relint-dim}(ii), which gives $\dim(\Scr_F(p')) \geq \dim(\spn(\simplexp - \{p\})) =  \dim(\affhull(\simplexp)) - d = \|p\|_0 - 1-d$.\footnote{To construct this second equality, observe $\dim(\spn(\simplexp - \{p\})) \leq \dim(\affhull(\simplexp))$ by using the standard basis for affine hull of the simplex.  For $\dim(\spn(\simplexp - \{p\})) \geq \dim(\affhull(\simplexp))$, observe that the uniform distribution on $\simplexp$ has full support and therefore requires $\|p\|_0 - 1$ elements in its basis.}

@@

\section{Omitted examples}\label{app:omitted-examples}
\paragraph{Discrete problem with no target loss.}
Consider the following scenario where someone is deciding how to dress for the weather based on a meteorologist's forecast.
Consider the three outcomes $\Y = \{$sunny, snowy, rainy$\}$, and we suppose we want to have some bias towards health and safety, so the meteorologist should only predict sunny weather if $Pr[$sunny $|$ weather data$] \geq 3/4$.
Otherwise, they should predict whatever is more likely  given the weather data: rain or snow.

We can now model this problem by a property with the reports $\R = \Y$, and have 
\begin{align*}
\gamma(p) &= \begin{cases}
\text{sunny} & p_{\text{sunny}} \geq 3/4 \\
\text{rainy} & p_{\text{sunny}} \leq 3/4 \wedge p_{\text{rainy}} \geq p_{\text{snowy}} \\
\text{snowy} & p_{\text{sunny}} \leq 3/4 \wedge p_{\text{snowy}} \geq p_{\text{rainy}} \\
\end{cases}~,~
\end{align*} 
shown in Figure~\ref{fig:t-example}.
Since the cells of elicitable properties in the simplex form a power diagram~\citep{lambert2009eliciting}, we know that there is actually \emph{no} target loss that directly elicits this problem.
Constructing a consistent surrogate for this task is ill-defined without Definition~\ref{def:consistent-prop}.
The function $\propdis(r,p) = \Ind{r \not \in \gamma(p)}$, which satisfies the requirements of $\propdis$, now allows us to use Definition~\ref{def:consistent-prop} to think about consistent surrogates for this task.

Intuitively, since the feasible subspace dimension bound would be lowest at the distribution $p = (1/8, 3/4,1/8)$, we might want to test Corollary~\ref{cor:Pcodim-flat-single-val-prop} or Corollary~\ref{cor:Pcodim-flat-elic-relint-prop} at $p$.
However, we cannot apply either at $p$ since $\gamma(p) = \{$rainy, snowy, sunny$\}$ but the property is not elicitable.
\citet[Theorem 16]{ramaswamy2016convex} cannot draw any conclusions about this property for two reasons that go hand in hand: first, we are given a target property instead of a target loss.
Second, since the property is not elicitable (hence why there can be no target loss), we observe $\dim(\Sc_{\gamma_{\text{rainy}}}(p)) \neq \dim(\Sc_{\gamma_{\text{sunny}}}(p))$, so their result cannot be applied as~\citet[Lemma 23]{ramaswamy2016convex} does not hold.

However, our bounds from Corollary~\ref{cor:Pcodim-flat-single-val-prop} on the distribution $q = (1/8, 3/4 - \epsilon, 1/8 + \epsilon)$ for a small enough $\epsilon > 0$, which we can apply since $\gamma(q) = \{$snowy$\}$, suggest that the convex elicitation complexity $\eliccvx(\gamma) \geq 2$, since there is no way to draw a $1$-flat (line) through $q$ while staying in just one level set on the simplex.

This example, although seemingly contrived, also extends to other decision-tree-like properties that do not have an explicit or easily constructed target loss.
\begin{figure}
	\centering
	\includegraphics[width=0.5\linewidth]{tikz/t-example.pdf}
	\caption{A meteorology example with a bias towards citizen safety.}
	\label{fig:t-example}
\end{figure}

>>>>

\section{Conclusions and future work}\label{sec:conclusions}
In this work, we show that indirect property elicitation can be a powerful necessary condition for the existence of a consistent surrogate loss (Theorem~\ref{thm:consistent-implies-indir-elic}).
Furthermore, we introduce a new lower bound (Corollaries~\ref{cor:Pcodim-flat-single-val-prop} and~\ref{cor:Pcodim-flat-elic-relint-prop}) on the dimension of a consistent convex loss that is generally applicable and extends previous results from both the discrete (Corollary~\ref{cor:fsd-bound}) and continuous (Corollaries~\ref{cor:variance} and \ref{cor:spectral-risks}) estimation settings.

@@

\newpage
\appendix
\section{A general notion of calibration}\label{app:calibration}
For general settings, we introduce a notion of calibration that is a special case of calibration as introduced by~\citet[Definition 2.7]{steinwart2007compare} and~\citet[Chapter 3]{steinwart2008support}. 
%\jessiet{I think we need to be careful how we present this; I'm not sure it's a new definition, though we seem to prove new results about it.}
We will show that in discrete prediction settings, it is equivalent to the more commonplace definition given in Definition~\ref{def:calibrated-finite}.
Therefore, we use this more general definition of calibration when proving statements about the relationship between consistency, calibration, and indirect elicitation.

\begin{definition}[Calibrated]\label{def:calibrated-general}
	A loss $L:\reals^d \times \Y \to \reals$ is \emph{$\P$-calibrated} with respect to a loss $\ell : \R \times \Y \to \reals$ eliciting the property $\gamma$ \jessie{$\gamma$ isn't used here...?} if there is a link $\psi : \reals^d \to \R$ such that, for all distributions $p \in \P$, there exists a function $\zeta : \reals_+ \to \reals_+$ with $\zeta$ continuous at $0^+$ and $\zeta(0) = 0$ such that for all $u \in \reals^d$, we have
	\begin{equation}\label{eq:calibrated-general}
	\ell( \psi(u); p) - \risk{\ell}(p)  \leq \zeta \left(  \exploss{L}{u}{p} - \risk{L}(p) \right)~.~
	\end{equation}
	If $\P = \simplex$, we simply say $(L, \psi)$ is calibrated.
\end{definition}

@@

\begin{proposition}
	When $\R$ and $\Y$ are finite, a continuous loss and link $(L, \psi)$ are $\P$-calibrated with respect to a target loss $\ell$ via Definition~\ref{def:calibrated-general} if and only if they are $\P$-calibrated via Definition~\ref{def:calibrated-finite}.
	\jessie{Change the order since we have forward refs?}
  %\btw{Okay if $\Gamma$ is empty}
\end{proposition}
\begin{proof}
$\implies$
	We prove the contrapositive; if $(L, \psi)$ is not calibrated with respect to $\ell$ by Definition~\ref{def:calibrated-finite}, then it is not calibrated via Definition~\ref{def:calibrated-general} either.
	If $(L, \psi)$ are not calibrated with respect to $\ell$ by Definition~\ref{def:calibrated-finite}, then there is a $p \in \P$ so that $\inf_{u : \psi(u) \not \in \gamma(p)} \exploss{L}{u}{p} = \inf_u \exploss{L}{u}{p}$.
	Thus there is a sequence $\{u_m\}$ so that $\lim_{m \to \infty} \psi(u_m) \not \in \gamma(p)$ and $\exploss{L}{u_m}{p} \to \risk{L}(p)$.  
	Now we have $R_L(u_m; p) \to 0$ but $R_\ell(\psi(u_m); p) \not \to 0$, so by Lemma~\ref{lem:calib-converging-regrets}, we contradict calibration by Definition~\ref{def:calibrated-general}.
%	
% Commented out 05.19.2020 for easier proof if Lemma 5 is true.
%	We prove the contrapositive; if $(L, \psi)$ is not calibrated with respect to $\ell$ by Definition~\ref{def:calibrated-finite}, then it is not calibrated via Definition~\ref{def:calibrated-general} either.
%	
%	Suppose there was a distribution $p \in \simplex$ so that $\inf_{u : \psi(u) \not \in \gamma(p)} L(u;p) = \inf_{u} L(u;p)$.
%	There must then be a sequence $\{u_m\} \to u$ so that $\lim_{m \to \infty} \psi(u_m) \not \in \gamma(p)$ and $L(u_m; p) \to \risk{L}(p)$.
%	
%	This consequently implies $R_L(u_m;p) \to 0$ as $L(u_m; p) \to \risk{L}(p)$, but as $\ell(\psi(u_m); p) \not \to \risk{\ell}(p)$ (if it did converge, then we would have $\psi(u_m) \to r \in \gamma(p)$), so $R_\ell(\psi(u_m); p) \not \to 0$.
%	Thus, by Lemma~\ref{lem:calib-converging-regrets}, we have no calibration via Definition~\ref{def:calibrated-general}.

$\impliedby$
Suppose there was a function $\zeta$ satisfying the bound in eq.~\eqref{eq:calibrated-general} for a fixed distribution $p \in \P$.
Observe the bound in eq.~\eqref{eq:calibration} can be written as $R_L(u,p) > 0$ for all $p \in \simplex$ and $u$ such that $\psi(u) \neq \gamma(p)$. 
By eq.~\eqref{eq:calibrated-general}, for any sequence $\{u_m\}$ so that $\psi(u_m) \not \to \gamma(p)$, we have must have $\zeta(R_\ell(\psi(u_m), p)) \not \to 0$ as we would otherwise contradict the bound in eq.~\eqref{eq:calibrated-general} since $R_\ell(\psi(u), p) \not \to 0$. 
Therefore $R_L(u_m, p) \not \to 0$; thus, the strict inequality holds.
\end{proof}

@@

Lemma~\ref{lem:calib-converging-regrets} now gives a condition equivalent to calibration without requiring one to already have a function $\zeta$ in mind.
\begin{lemma}\label{lem:calib-converging-regrets}
	A continuous surrogate and link $(L,\psi)$ are $\P$-calibrated (via definition~\ref{def:calibrated-general}) with respect to $\ell$ if and only if, for all $p \in \P$ and sequences $\{u_m\}$ so that $R_L(u_m; p) \to 0$, we have $R_\ell(\psi(u_m); p) \to 0$.
	\jessie{$(A \wedge C) \iff D$}
\end{lemma}
\begin{proof}
\jessie{$(A \wedge C) \implies D$}
	$\implies$ Take a sequence $\{u_m\}$ so that $R_L(u_m;p) \to 0$.
	Since $\zeta(0) = 0$ and $\zeta$ is continuous at $0$, we have $\zeta(R_L(u_m;p)) \to 0$.
	As the bound from Equation~\eqref{eq:calibrated-general} is satisfied for all $u \in \reals^d$ by assumption, we observe
	\begin{align*}
	\forall m, \; &0 \leq R_\ell(\psi(u_m); p) \leq \zeta(R_L(u_m;p))\\
	\implies &0 \leq \lim_{m \to \infty} R_\ell(\psi(u_m); p) \leq \lim_{m \to \infty} \zeta(R_L(u_m;p)) = 0\\
	\implies &0 = \lim_{m\to\infty} R_\ell(\psi(u_m); p) ~.~
	\end{align*}

@@

  (ii)
  We first show $\spn(F_p) = \Scr_F(p)$.
  First, take $v \in \Scr_F(p)$, and take $\epsilon_0$ as in the definition.
  For $\epsilon = \epsilon_0 / 2$, we then have $p + \epsilon v \in F \implies \epsilon v \in F_p$, and therefore, $v \in \spn(F_p)$.  
  Now take $v \in \spn(F_p)$.
  Since $p \in \relint{F}$ (i), we have $\vec 0 \in \relint{F_p}$.
  Therefore there is an $\epsilon_0 > 0$ so that $\epsilon v \in F_p$ for all $\epsilon \in (-\epsilon_0, \epsilon_0)$ by convexity of $F$.
  Therefore, $v \in \Scr_F(p)$, and we observe $\Scr_F(p) = \spn(F_p)$.

@@

	Taking $\P = \simplexp$, we know $p' \in \relint{\simplexp}$ by construction, so we can apply Lemma~\ref{lem:finite-relint-dim}(ii), which gives $\dim(\Scr_F(p')) \geq \dim(\spn(\simplexp - \{p\})) =  \dim(\affhull(\simplexp)) - d = \|p\|_0 - 1-d$.\footnote{To construct this second equality, observe $\dim(\spn(\simplexp - \{p\})) \leq \dim(\affhull(\simplexp))$ by using the standard basis for affine hull of the simplex, which has $\|p\|_0 - 1$ elements.  For $\dim(\spn(\simplexp - \{p\})) \geq \dim(\affhull(\simplexp))$, observe that the uniform distribution on $\simplexp$ has full support and therefore requires $\|p\|_0 - 1$ elements in its basis.}

@@

\section{Omitted examples}\label{app:omitted-examples}
\paragraph{Discrete problem with no target loss.}
Consider the following scenario where someone is deciding how to dress for the weather based on a meteorologist's forecast.
Consider the three outcomes $\Y = \{$rainy, sunny, snowy$\}$, and we suppose we want to have some bias towards health and safety, so the meteorologist should only predict sunny weather if $Pr[$sunny $|$ weather data$] \geq 3/4$.
Otherwise, they should predict whatever is more likely  given the weather data: rain or snow.

We can now model this problem by a property with the reports $\R = \Y$, and have 
\begin{align*}
\gamma(p) &= \begin{cases}
\text{sunny} & p_{\text{sunny}} \geq 3/4 \\
\text{rainy} & p_{\text{sunny}} \leq 3/4 \wedge p_{\text{rainy}} \geq p_{\text{snowy}} \\
\text{snowy} & p_{\text{sunny}} \leq 3/4 \wedge p_{\text{snowy}} \geq p_{\text{rainy}} \\
\end{cases}~,~
\end{align*} 
shown in Figure~\ref{fig:t-example}.
Since the cells of elicitable properties in the simplex form a power diagram~\citep{lambert2009eliciting}, we know that there is actually \emph{no} target loss that directly elicits this problem.
Constructing a consistent surrogate for this task is ill-defined without Definition~\ref{def:consistent-prop}.
The function $\propdis(r,p) = \Ind{r \not \in \gamma(p)}$ 
%, which satisfies the requirements of $\propdis$, 
now allows us to use Definition~\ref{def:consistent-prop} to think about consistent surrogates for this task.

Intuitively, since the feasible subspace dimension bound would be lowest at the distribution $p = (1/8, 3/4,1/8)$, we might want to test Corollary~\ref{cor:Pcodim-flat-single-val-prop} or Corollary~\ref{cor:Pcodim-flat-elic-relint-prop} at $p$.
However, we cannot apply either at $p$ since $\gamma(p) = \{$rainy, snowy, sunny$\}$ but the property is not elicitable.
\citet[Theorem 16]{ramaswamy2016convex} cannot draw any conclusions about this property for two reasons: first, we are given a target property instead of a target loss.
Second, since the property is not elicitable (hence why there can be no target loss), we observe $\dim(\Sc_{\gamma_{\text{rainy}}}(p)) \neq \dim(\Sc_{\gamma_{\text{sunny}}}(p))$, contradicting the requirements of~\citet[Lemma 23]{ramaswamy2016convex}.

However, our bounds from Corollary~\ref{cor:Pcodim-flat-single-val-prop} on the distribution $q = (1/8, 3/4 - \epsilon, 1/8 + \epsilon)$ for a small enough $\epsilon > 0$, which we can apply since $\gamma(q) = \{$snowy$\}$, suggest that the convex elicitation complexity $\eliccvx(\gamma) \geq 2$, since there is no way to draw a $1$-flat (a line, since $q \in \relint{\simplex}$) through $q$ while staying in just one level set on the simplex.

This example also extends to other decision-tree-like properties that do not have an explicit or easily constructed target loss.
\begin{figure}
	\centering
	\includegraphics[width=0.3\linewidth]{tikz/t-example.pdf}
	\caption{A meteorology example with a bias towards citizen safety.}
	\label{fig:t-example}
\end{figure}
